Mainstreaming Climate Adaptation Requires Subjective Integration of Deep Uncertainties

### 2022-02-26 papers/paper/preprint.tex
some edits by JDG, going to send to Klaus for feedback

% fonts and appearance
\usepackage{amsmath, amsfonts, physics, siunitx}
\usepackage[american]{babel} 
\usepackage[T1]{fontenc} % improved font encoding
\usepackage[ttscale=0.8]{libertine}
\usepackage{fontawesome5}
\usepackage[format=plain, textfont=it]{caption}

% page size and margins
\usepackage{geometry}
\geometry{letterpaper,top=1in, bottom=1in, left=0.5in, right=2in}

@@

% ACRONYMS
\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
\newacronym{bdt}{BDT}{Bayesian decision theory}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym[]{cdf}{CDF}{cumulative distribution function}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym[]{gev}{GEV}{generalized extreme value}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{lsl}{LSL}{local mean sea level}
\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{slr}{SLR}{relative sea level rise}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
\newacronym{tc}{TC}{tropical cyclone}

@@

\begin{abstract}
    The policy and engineering tools used to inform climate adaptation etc rely on standards and cost-benefit analyses that depend on risk probabilities.
    One example is house elevation: the recommended elevation is a standard (BFE plus a foot) and funding decisions rely on CBA.
    Under stationarity, the probabilities for this can almost be agreed upon.
    Under nonstationarity, there are deep uncertainties.
    We show that the way these deep uncertainties are synthesized into CBA matters a lot.
    Thus, static decisions are subject to deep uncertainties
\end{abstract}

\clearpage

\section{Introduction}\label{sec:introduction}

Since infrastructure built today is likely to persist for several decades, infrastructure designs must take into account the stresses that a project is likely to face decades into the future.
On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.\james{Too fast on intro. Need to slow down and define concepts}
On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
For example, \ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}, something by Flyvbjerg, \etc}

TODO: set up flood risk management as a context where over- or under-design could be relevant, then raise house elevation in the coastal zone as a specific example.
Federal guidance for this decision relies on standards (the $T$-year flood) rather than risk-based design.
Be sure to note that this has parallels to lots of other engineering design guidance. Then note the work of \citet{zarekarizi_suboptimal:2020} or \citet{xian_elevation:2017}, who show that a risk-based approach may improve considerably on a standards-based approach.

Additionally, the floodplain maps used for decision supportare silent on projected future changes and neglect the deep and dynamic uncertainties surrounding the projected hazards\ldots
Both risk-based design and standards-based approaches are complicated by nonstationarity.
On the one hand, nonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{slr} is clear.
On the other, efforts to find objective methods for quantifying nonstationarity, eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification \citep[\eg,][]{keller_management:2021}.

This motivates subjective approaches for designing for nonstationarity.
Thus decision-makers need some subjective way to choose which possible futures to design for.

In this paper, we use a didactic case study of house elevation in Norfolk, VA as a lens through which to shed light on the following research questions.
\begin{description}
    \item[Q1] What are the potential benefits to homeowners of changing threshold-based guidance to risk-based guidance under nonstationary flood hazard?
    \item[Q2] How does risk-based guidance change under different assumptions of physical model structure and human emissions forcings?
    \item[Q3] How can deep uncertainties be transparently and consistently synthesized for decision analysis?
\end{description}
Specifically, we examine this problem through three lenses.\james{Which ones?}

The paper is organized as follows.\james{This is confusing to the reader}
In \cref{sec:case-study} we describe our case study in Norfolk, VA.
First, in \cref{sec:multiple-simulation} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
Next, in \cref{sec:multiple-pdf} we consider the design problem as a set of ``multiple \glspl{pdf}'' and use these findings to motivate theoretical advances.
Then in \cref{sec:synthesizing} we present a probabilistic method for synthesizing simulations from multiple \glspl{pdf} and illustrate the advantages and potential pitfalls of this approach.
Finally in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify  future research needs.

@@

We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard.
For interpretability, we focus on deep uncertainty in \gls{slr} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
We consider a single, one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
We evaluate this decision for each of many possible simulations of \gls{slr} using a combined engineering-economic model based on that of \citet{zarekarizi_suboptimal:2020}.
In the remainder of this section we discuss data sources and the system model.

\begin{table}
    \centering
    \caption{
        Summary of parameters, their notation, and how their uncertainty is modeled.
    }\label{tab:uncertainties}
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name             & Symbol           & Uncertainty                                                                          & See more                     \\
        \midrule
        \Gls{slr}        & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \cref{sec:sea-level}         \\
        Storm surge      & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \cref{sec:storm-surge}       \\
        Discount rate    & $1-\gamma$       & Fixed at 3\%                                                                         & \cref{sec:led}               \\
        Depth-damage     & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
        Elevation cost   & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \cref{fig:cost-up-front}     \\
        Initial height   & $h_0$            & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 &                              \\
        House floor area &                  & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                &                              \\
        Structural value &                  & Deterministic: \usd{200000} unless otherwise noted                                   &                              \\
        House lifetime   &                  & Deterministic: 70 years unless otherwise noted                                       &                              \\
        \bottomrule
    \end{tabular}
\end{table}

@@

Deeply uncertain sea level rise is the focus of our model.
Our analysis considers the common case where a decision maker or analyst has access to an existing set of simulations.
Specifically, we use simulations published in \citet{ruckart_hazardous:2008} that use the BRICK model version 0.2 \citep{wong_brick0.2:2017}.

\Cref{fig:lsl-evolution} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways \citep[\gls{rcp} scenarios][]{moss_rcp:2010} and two models of ice sheet dynamics.
Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
This is also the \gls{pdf} with the greatest within-model variance.
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lsl-evolution}
    \caption{
        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for three probabilistic models.
        (B): probability distribution of mean sea level at Sewells Point, VA in the year 2100 for all probabilistic models considered.
    }\label{fig:lsl-evolution}
\end{figure}

For consistency with previous studies, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \gls{sow}}.

@@

Neglecting hydrodynamics, we model annual maximum floods $y_t$ as the sum of sea level $\overline{y}_t$ and storm surges $y'_t$.

Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).\james{cite NOAA data source}
Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
For each year we remove the annual mean then extract a single maximum; we refer the set of these maxima for all years as the time series of annual maximum storm surges.
This time series is shown in \cref{fig:surge-obs-return}(a).
In (b) we use the Weibull plotting position
\begin{equation}\label{eq:plot-pos}
    \frac{r}{N + 1}
\end{equation}
where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{surge-obs-return}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
        (A):
        Time series of historic storms.
        Blue (green) arrows denote notable tropical cyclones (nor'easters).
        (B):
        Return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit.
    }\label{fig:surge-obs-return}
\end{figure}

We model future storm surge risk through a probabilistic lens.
To do this, we develop a stationary\james{discuss / validate this -- it's cool, we have the supplemental info and can emphpasize our focus here is on SLR for didactic purposes} Bayesian model for storm surge:
\begin{equation}\label{eq:surge-model}
    y'_t \sim \text{GEV}\left(\mu, \sigma, \xi \right),
\end{equation}
where $y'_t$ is the storm surge (above sea level) in year $t$ and $\mu$, $\sigma$, and $\xi$ are the location, scale, and shape parameters, respectively.

When \gls{gev} distributions are fit to time series data without strong prior information, the resulting uncertainties can be unrealistically large.\james{cite}
Thus, information from other locations, time periods, and/or from physical principles may help to reduce uncertainties \citep{Merz:2008bl,Merz:2008eh}.
To constrain estimates, we add weakly informative priors.
However, rather than adding prior information over the parameters themselves, we follow \citet{coles_evd:1996} and apply a prior over extreme quantiles of the distribution.
Specifically, we apply LogNormal priors over the 90th, 99th, and 99.9th percentiles of the distribution (\ie, the 10, 100, and 1000 year return period events).\james{Expand on this ``probabilistic inversion''}
These values were chosen to represent plausible physical ranges (the assumed 90\% confidence interval for the 10 year event is XY years, for the 100 year event is XZ years, and for the 1000 year event ZY years) and are plotted in supporting figure SX.\james{return to this sentence}

To evaluate a single \gls{sow}, we integrate over $J$ samples from the posterior distribution (of \cref{eq:surge-model}).

@@

There are two components of the system model (``relationships'' in \cref{fig:xlrm}).
The first is a fragility model that estimates the expected flood loss for a particular year, given the elevation of the house and the mean sea level for that year.
Specifically, expected annual damages in year $t$ is defined as the expected damages in a given year, conditional on the house's height relative to the gauge after elevation ($h = h_0 + \Delta h$):
\begin{equation}\label{eq:ead}
    \textrm{EAD}_t = \mathbb{E}[d_t | h] = \int_{y_t} p(y_t) D(h - y_t) dy_t,
\end{equation}
where $y_t = \overline{y}_t + y'_t$ is the annual maximum flood, relative to the gauge, in year $t$, and $D(h - y_t)$ is the damage as a function of depth relative to the house.
Following \citet{zarekarizi_suboptimal:2020} we use the HAZUS deterministic depth-damage relationship to parameterize $D(h-y_t)$ (see fig. XYZ(c)).\james{Cite HAZUS}

The expected annual damages is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method:
\begin{enumerate}
    \item For $k=1, \ldots, K$:
          \begin{enumerate}
              \item sample a draw from the posterior distribution of flood hazard to get $\left\{ \mu_k, \sigma_k, \xi_k \right\}$
              \item simulate a single storm surge from the stationary distribution (equation XYZ) and add the mean sea level to get total flood depth
              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_t$) into  the deterministic HAZUS depth-damage relationship
              \item store this as the $k$th damage
          \end{enumerate}
    \item Estimate expected annual damages as the sample mean of the $K$ estimates
\end{enumerate}

However, evaluating expected annual damages for each of $N$ simulations of sea level rise, each of $J$ draw from the posterior distribution of storm surge, and each of $T$ time steps requires $N \times J \times T \times K$ simulations.
Since $K$ must be large in order to accurately approximate the integral, this requires incurring a heavy computational cost.
However, noticing that this function dependsonly on the elevation of the house relative to \gls{slr}, we develop a simuple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
To do this, we  precompute expected annual damage for all height differences in \SI{0.1}{ft} increments from -XY to YX ft and fit a piecewise linear interpolation to this data.\james{check increements, piecewise linear}
We use $K=\num{1e6}$ samples to fit this emulator for each of the NUMBER increments.
This model is shown in supplemental figure SX.\james{supplemental figure}
Once this interpolation has been precomputed, calculating expected annual damage for a particular year only requires evaluating a piecewise linear function.

>>>>

% fonts and appearance
\usepackage{amsmath, amsfonts, physics, siunitx, nicefrac}
\usepackage[american]{babel} 
\usepackage[T1]{fontenc} % improved font encoding
\usepackage[ttscale=0.8]{libertine}
\usepackage{fontawesome5}
\usepackage[format=plain, textfont=it]{caption}

% page size and margins
\usepackage{geometry}
\geometry{letterpaper,top=1in, bottom=1in, left=1in, right=2in}

@@

% ACRONYMS
\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
\newacronym{bdt}{BDT}{Bayesian decision theory}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym[]{cdf}{CDF}{cumulative distribution function}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym[]{gev}{GEV}{generalized extreme value}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{lsl}{LSL}{local mean sea level}
\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{msl}{MSL}{mean relative sea level}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
\newacronym{tc}{TC}{tropical cyclone}

@@

\begin{abstract}
    National agencies, regional governments, and individual households manage flood risk through a combination of standards and cost-benefit analyses that rely on estimates of flood hazard.
    One example is house elevation: \acrlong{fema} recommends a house located in the floodplain be elevated to the \acrlong{bfe} (typically the 100 year flood level) plus a freeboard (typically \SI{1}{ft}) if doing so would pass a cost-benefit test.
    Under stationarity, the probabilities used for this analysis (\ie to estimate the 100 year flood height or to perform the cost-benefit analysis) can be estimated using empirical frequencies from the observational record.
    However, growing recognition of nonstationarity motivates a forward-looking approach.
    Efforts to incorporate nonstationarity hazard into planning have struggled to overcome deep uncertainties in projections of future hazard because different models  may fit the observed data equally well but yield very different estimates of future hazard, and there is often no objective way to choose between them (\ie, equifinality).
    In this paper we use a didactic case study of deciding if and how high to elevate a single house in the coastal zone to illustrate a decision theoretic framework that combines exploratory modeling, robust decision making, and subjective Bayesian philosophy.
    We first use exploratory modeling to illustrate that there is no dominant strategy; instead, there are instrinsic tradeoffs between performance under low- and high-sea level rise scenarios.
    Next, we map the tradeoff between up-front cost and expected damages for each of sixteen models of future sea level rise to illustrate that the choice of which model to use has a strong influence on estimated trade-offs, cost-benefit analyses, or policy prescriptions.
    To overcome these shortcomings, we propose and illustrate a computationally efficient method for scenario weighting conditional on a subjective prior belief over future sea level rise.
    This method provides a transparent framework for synthesizing deep uncertainties, facilitating critique and dialogue around necessarily subjective decisions.
    It can be applied beyond nontationary flood risk management to other problems in climate adaptation, calculating the social cost of carbon, and more.
\end{abstract}

Key points
\begin{enumerate}
    \item
\end{enumerate}

\clearpage
\section{Introduction}\label{sec:introduction}

Estimates of flood hazard affect peoples' lives
    \item Floods affect lots of people and property
    \item Climate change and exposure making floods worse
    \item Floodplain management: $T$ year flood
    \item Some cost-benefit analysis and lots of design standards / heuristics
One example is floodproofing
    \item Literature on floodproofing and building-scale vulnerability reduction measures
    \item This works great!
    \item But who should use it? We're still using standards.
    \item Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
          Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
          For example, \citet{xian_elevation:2017} used a cost-benefit analysis to demonstrate that tailoring recommendations to the inital elevation of a structure can reduce expected costs.
          \citet{zarekarizi_suboptimal:2020} show that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
          Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
Nonstationarity
    \item These analyses all rely on \glspl{pdf}
    \item Growing evidence for nonstationarity in a wide variety of hydroclimate processes
    \item We often focus on climate change as a driver but sometimes local modifications matter just as much \citep{Merz:2014gf}
    \item Even where risk-based analyses are used, this problem recommendations
    \item Yet stationarity remains the dominant assumption because it is immortal \citep{Montanari:2014hl} or undead \citep{Serinaldi:2015bq}
    \item Bonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{msl} is clear.
    \item On the other, efforts to find objective methods for quantifying nonstationarity, eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification
It matters how you prepare
    \item Equifinality problem: different models diverge even if they fit the historical record well
    \item Over- and under-design can each be costly \citep{DossGollin:2019}
    \item On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
          For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.
    \item On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future. For example,\ldots\ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}, something by Flyvbjerg, \etc}
A quick philosophical literature review of how people have approached and interpreted this problem
    \item This motivates subjective approaches for designing for nonstationarity.
    \item There's a little bit of RDM-style work on floods \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}
    \item Model choice in $\mathcal{M}$-open case
    \item Philosophy of Bayesian statistics and decision making -- explain subjective probability
Case study
    \item House elevation in the coastal zone as a specific example of a problem where both over- or under-investment could be a problem
    \item Federal guidance uses both standards (floodplain, \gls{bfe}) and cost-benefit analysis (does it pass test?)
    \item Be sure to note that this has parallels to lots of other engineering design guidance.
    \item \citet{xian_elevation:2017}: one size does not fit all
    \item \citet{zarekarizi_suboptimal:2020}: neglecting uncertainty leads to bad outcomes
Plan d'attaque
    \item Research questions
          \begin{description}
              \item[Q1] What are the potential benefits to homeowners of changing threshold-based guidance to risk-based guidance under nonstationary flood hazard?
              \item[Q2] How does risk-based guidance change under different assumptions of physical model structure and human emissions forcings?
              \item[Q3] How can deep uncertainties be transparently and consistently synthesized for decision analysis?
          \end{description}
    \item In \cref{sec:case-study} we describe our case study in Norfolk, VA.
    \item  in \cref{sec:multiple-simulation} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
    \item  in \cref{sec:multiple-pdf} we consider the design problem as a set of ``multiple \glspl{pdf}'' and use these findings to motivate theoretical advances.
    \item  in \cref{sec:synthesizing} we present a probabilistic method for synthesizing simulations from multiple \glspl{pdf} and illustrate the advantages and potential pitfalls of this approach.
    \item  in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify  future research needs.

@@

We model a single, one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
We evaluate this decision for each of many possible simulations of \gls{msl} using a combined engineering-economic model based on that of \citet{zarekarizi_suboptimal:2020}.
In the remainder of this section we discuss data sources and models.

\begin{table}
    \centering
    \caption{
        Summary of parameters, their notation, and how their uncertainty is modeled.
    }\label{tab:uncertainties}
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name             & Symbol            & Uncertainty                                                                          & See more                     \\
        \midrule
        \Gls{msl}        & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        & \cref{sec:sea-level}         \\
        Storm surge      & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               & \cref{sec:storm-surge}       \\
        Ann-max flood    & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          &                              \\
        Discount rate    & $1-\gamma$        & Fixed at 3\%                                                                         & \cref{sec:led}               \\
        Depth-damage     & $D(h-y)$          & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
        Elevation cost   & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \cref{fig:cost-up-front}     \\
        Initial height   & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 &                              \\
        House floor area &                   & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                &                              \\
        Structural value &                   & Deterministic: \usd{200000} unless otherwise noted                                   &                              \\
        House lifetime   &                   & Deterministic: 70 years unless otherwise noted                                       &                              \\
        \bottomrule
    \end{tabular}
\end{table}

@@

Deeply uncertain sea level rise is the dominant source of deep uncertainty in our analysis.
Decision makers and analysts often have access to simulations of physical variables like \gls{msl}, but usefully incorporating these simulations into decision analyses remains an area of active research.

To inform the house elevation decision, we develop an ensemble of simulations of future \gls{msl}.
Specifically, we analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic models using data published in \citet{ruckert_coastal:2019}.
The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
The \citet{kopp_probabilistic:2014} and \citet{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
These four models paramaterize physical processes, particularly of ice sheet dynamics, in different ways.
For a discussion of these methods, their similarities, their differences, and their consistency with other analyses see \citet{ruckert_coastal:2019}, \citet{kopp_evolving:2017}, or \citet{bamber_slrise:2019}.

Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain.
To represent this uncertainty, we sample four \gls{rcp} scenarios.
This gives us sixteen time-varying probability density functions of \gls{msl}.
We refer to each time series of \gls{msl}, $\overline{y}(t)$, as \emph{a \gls{sow}} and to each combination of physical model and \gls{rcp} scenario as \emph{a model} for $p(\overline{y}|t)$.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lsl-evolution}
    \caption{
        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for a representative subset of three probabilistic models (out of sixteen).
        (B): probability distribution of \gls{msl} at Sewells Point, VA in the year 2100 for each probabilistic model considered.
    }\label{fig:lsl-evolution}
\end{figure}

\Cref{fig:lsl-evolution}(a) shows the time-varying \glspl{pdf} of \gls{msl} for three representative models.
The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
Both physical model and \gls{rcp} scenario strongly inform the modeled probability distribution.\klaus{Do you want to add a sentence here comparing BRICK to the Kopp et al models?}

@@

Coastal flood hazard does not depend only on \gls{msl} but also on storm surges.
Neglecting hydrodynamics, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$ and annual maximum storm surges $y'(t)$.
In this subsection we describe our representation of storm surge.

Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).
The NOAA tides and currents dataset is available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
For each year we first remove the annual mean, then extract the  maximum water level.
We refer these maxima as the annual maximum storm surges $y'(t)$.

\Cref{fig:surge-obs-return}(a) shows the time series of historical annual maximum storm surges.
Major historical surges have been driven by two different mechanisms: tropical cyclones and nor'easters.
The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and nor'easters have caused surges above \SI{6}{ft}.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{surge-obs-return}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
        (A):
        time series of historic storms.
        Red (yellow) arrows denote notable tropical cyclones (nor'easters).
        (B):
        return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\ref{eq:plot-pos}).
        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:storm-surge}).
    }\label{fig:surge-obs-return}
\end{figure}

We model future storm surge using a stationary \gls{gev} model:
\begin{equation}\label{eq:surge-model}
    y'(t) \sim \text{GEV}\left(\mu, \sigma, \xi \right),
\end{equation}
where $y'(t)$ is the storm surge (above sea level) in year $t$ and $\mu$, $\sigma$, and $\xi$ are the location, scale, and shape parameters, respectively, of the \gls{gev} distribution.
This model is stationary because there is no time dependence; surges in different years are modeled as \gls{iid}.

To improve estimates, we add twp forms of prior information.
First, we constrain $\xi > 0$ to reflect knowledge about the support of $y'$: when $\xi < 0$, $y' \in \qty(-\infty, \mu - \nicefrac{\sigma}{\xi})$  and when $\xi > 0$, $y' \in \qty(\mu-\nicefrac{\sigma}{\xi}, \infty)$.
Since storm surges cannot be negative, only the latter is physically defensible.
Second, we add weakly informative priors.
Rather than applying prior information directly over the joint distribution of the priors $\qty{\mu,\sigma,\xi}$, for which intuition is difficult, we apply a prior over extreme quantiles of the distribution as in \citet{coles_evd:1996}.
Specifically we apply half Normal (\ie, Normal distributions truncated at zero) priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
These values were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.

For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \citep{Betancourt:2017vd,hoffman_nuts:2011} through the Turing package of the Julia programming language \citep{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
We evaluate the model's fit using posterior predictive checks \citep[see section 2.4 of][and references therein]{gelman_workflow:2020}.
Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesisan test statistics, we found that draws from the posterior predictive distribution matched the observed test statistic credibly \cref{fig:surge-test-statistics}.
Since \cref{fig:surge-test-statistics}(b) suggests the possibility of meaningful temporal structure not captured by our stationary \gls{iid} model, future efforts  could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020}.

\Cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges.
The return period of the data (dots) is shown using the Weibull (``empirical'') plotting position
\begin{equation}\label{eq:plot-pos}
    \frac{r}{N + 1}
\end{equation}
where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
The good fit of the Bayesian fit (gray shading) to the data (points) suggests that the stationary approximation is reasonable for this dataset.

@@

There are two components of the system model (``relationships'' in \cref{fig:xlrm}).
The first is a fragility model that estimates the expected flood loss for a particular year, given the elevation of the house and the mean sea level for that year.
Specifically, we define expected annual damages in year $t$ as the expectation of the damage function with respect to storm surge.
This expectation depends on the house's height ($h = h_0 + \Delta h$) where $h_0$ is the initial height relative to the gauge and $\Delta h$ is the amount by which the house is elevated.
The expcted annual damage is thus
\begin{equation}\label{eq:ead}
    \textrm{EAD} = \mathbb{E}[D(h-\overline{y})] = \int_{y'} p(y') D(h - (\overline{y} + y')) \dd{y'}
\end{equation}
where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
The term $\overline{y}$ in \cref{eq:ead} is time-dependent but we have omitted the time dependence for readability.
Following \citet{zarekarizi_suboptimal:2020} we use the we use the Hazard U.S. (HAZUS) depth-damage curves provided by FEMA to parameterize $D(h-\overline{y})$; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
For comparison, \cref{fig:cost-depth-damage} also shows results using the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
Although \citet{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for concision.
